---
title: "Bicycle-Theft-Report"
author: "Chris Yan 1009096746, Aaron Dong 1008961232, Armaan Rehman Shah 1009641309"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Importing the dataset
origin <- read.csv("bicycle.csv")
```


```{r, echo=FALSE}
# Initiating libraries
library(tidyverse)
library(ggplot2)
library(ggmap)
library(png)
library(ggpubr)
library(osmdata)
citation("ggmap")
```
The report concerns with the data "Bicycle Thefts Open Data" which contains occurrences related to bicycle thefts reported to the Toronto Police Service. These occurrences are related to a variety of offences where the theft of a bicycle was included.
\
\
**Dataset Description**
* entry id named X_id
  * unique entry id in increasing order
* event id named event_unique_id
  * unique event id not sorted
* offense committed named Primary_Offense
  * type of crime that is committed
* offense occurrence date named Occurrence_Date
  * date of when the crime was committed
* offense occurrence year named Occurrence_Year
  * year of when the crime was committed
* offense occurrence month named Occurrence_Month
  * month of when the crime was committed
* offense occurrence day of the week named Occurrence_DayOfWeek
  * what days of week was the crime committed (from Mon to Sun)
* offense occurrence date of the month named Occurrence_DayOfMonth
  * what date of the month was the crime committed (from 1-31)
* offense occurrence day of the year named Occurrence_DayOfYear
  * what day of the year was the crime committed (from 1-365)
* offense occurrence hour named Occurrence_Hour
  * what hour of the day was the crime committed (from 1-24)
* offense report date named Report_Date
  * when the incident was reported to the authorities
\
\
**Purpose:** The purpose of this report is to study the ...

```{r}
#omit all the entries that has NA inputs, they won't be helpful to our data
data <- na.omit(origin)
summary <- table(data$Occurrence_Year)
print(summary)
```

We can see that there are some outliers from 2009-2013, in addition, the 2022 data is not complete.
Therefore, Let's filter them out and obtain a complete set of data, the one we will be using for our report below.
```{r}



```
