---
title: "Bicycle-Theft-Report"
author: "Chris Yan 1009096746, Aaron Dong 1008961232, Armaan Rehman Shah 1009641309"
date: "2023-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
# Importing the dataset
origin <- read.csv("bicycle.csv")
```


```{r, echo=FALSE, eval=FALSE}
# Initiating libraries
library(tidyverse)
library(ggplot2)
library(ggmap)
library(png)
library(ggpubr)
library(osmdata)
```

## Introduction

The report concerns with the data "Bicycle Thefts Open Data" which contains occurrences related to bicycle thefts reported to the Toronto Police Service. These occurrences are related to a variety of offences where the theft of a bicycle was included. We obtained this dataset from the City of Toronto Open Data portal. It records all of the bike thefts in Toronto starting from as early as 2013. The dataset is published by the Toronto Police services and is refreshed annually all the way up until 2022. The dataset contains large amount of entries and multiple information variables.

### Dataset Description

* __entry id named X_id__
  * unique entry id in increasing order
  * _integer_
* __event id named event_unique_id__
  * unique event id not sorted
  * _string_
* __offense committed named Primary_Offense__
  * type of crime that is committed
  * _string_
* __offense occurrence date named Occurrence_Date__
  * date of when the crime was committed
  * _string_
* __offense occurrence year named Occurrence_Year__
  * year of when the crime was committed
  * _integer_
* __offense occurrence month named Occurrence_Month__
  * month of when the crime was committed
  * _string_
* __offense occurrence day of the week named Occurrence_DayOfWeek__
  * what days of week was the crime committed (from Mon to Sun)
  * _string_
* __offense occurrence date of the month named Occurrence_DayOfMonth__
  * what date of the month was the crime committed (from 1-31)
  * _integer_
* __offense occurrence day of the year named Occurrence_DayOfYear__
  * what day of the year was the crime committed (from 1-365)
  * _integer_
* __offense occurrence hour named Occurrence_Hour__
  * what hour of the day was the crime committed (from 1-24)
  * _integer_
* __offense report date named Report_Date__
  * when the incident was reported to the authorities
  * _string_
* __offense report year named Report_Year__
  * which year was the incident reported
  * _integer_
* __offense report month named Report_Month__
  * which month was the incident reported
  * _string_
* __offense report day of the week named DayOfWeek__
  * which day of week was the incident reported (Mon - Sun)
  * _string_
* __offense report day of the month named DayOfMonth__
  * which day of the month was the incident reported (from 1-31)
  * _integer_
* __offense report day of the year named DayOfYear__
  * which day of the year was the incident reported (from 1-365)
  * _integer_
* __offense report hour named Report_Hour__
  * which hour of the day was the incident reported
  * _integer_
* __division where offense named Division__
  * the division of the location, mostly used to identify the police service responsible
  * _string_
* __city where offense where offense took place named City__
  * city name in or around Toronto
  * _string_
* __hood where offense took place named Hood_ID__
  * the hood id for all Toronto's neighborhoods (out of 158)
  * _string_
* __neighborhood where offense took place named NeighbourhoodName__
  * the neighborhood name in Toronto and surrounding areas
  * _string_
* __location where offense took place named Location_Type__
  * broad description of the crime location.
  * _string_
* __location premise of where offense took place named Premise_Type__
  * description of where bike was stolen
  * _string_
* __bike make of stolen bike named Bike_Make__
  * brand of bike that was stolen.
  * _string_
* __bike model of stolen bike named Bike_Model__
  * model of bike that was stolen.
  * _string_
* __bike type of stolen bike named Bike_Type__
  * type of bike that was stolen
  * _string_
* __bike speed of stolen bike named Bike_Speed__
  * the assumed speed of the bike
  * _integer_
* __bike color of stolen bike named Bike_Colour__
  * colour of the bike that was stolen.
  * _string_
* __cost of stolen bike named Cost_of_Bike__
  * cost of bike that was stolen.
  * _double_
* __condition of stolen bike named Status__
  * current condition of the bike (STOLEN, RECOVERED, or UNKNOWN)
  * _string_
* __coordinates of bike stolen named Geometry__
  * precise location of the crime location
  * _string_
\
\
**Purpose:** The purpose of this report is to study the ...

```{r}
#omit all the entries that has NA inputs, they won't be helpful to our data
data <- na.omit(origin)
summary <- table(data$Occurrence_Year)
print(summary)
```

We can see that there are some outliers from 2009-2013, in addition, the 2022 data is not complete.
Therefore, Let's filter them out and obtain a complete set of data, the one we will be using for our report below.
```{r}



```
